\RequirePackage[utf8]{inputenc} % remove when using lualatex oder xelatex!
\RequirePackage{hgbpdfa}        % Creates a PDF/A-2b compliant document

\documentclass[english]{hgbarticle}

%%%----------------------------------------------------------
\begin{document}
%%%----------------------------------------------------------

\title{The \textsf{hagenberg-thesis} Package}
\date{\hgbDate}

\author{W.\ Burger and W.\ Hochleitner\\[10pt]
University of Applied Sciences Upper Austria\\
Department of Digital Media, Hagenberg (Austria)}

%%%----------------------------------------------------------
\maketitle
%%%----------------------------------------------------------

\begin{abstract}\noindent
The \textsf{hagenberg-thesis} package is a collection of modern LaTeX
templates for university theses (bachelor, master, or diploma programs) and
related documents. This manual describes the main features of this package.
Pre-configured document templates for English and German manuscripts and a
complete tutorial are available on the package's home repository.
\end{abstract}


\section{Introduction}

The complete source of this package and auxiliary materials are available on
CTAN%
\footnote{\url{https://ctan.org/pkg/hagenberg-thesis}}
and its development repository.%
\footnote{\url{https://github.com/Digital-Media/HagenbergThesis}}
The package is made available under the terms of the Creative Commons
Attribution 4.0 International Public License.%
\footnote{\url{https://creativecommons.org/licenses/by/4.0/legalcode}}


\section{Document classes}

The \texttt{hgb} package provides the following document classes, which are
based on the standard \latex\ classes \texttt{book}, \texttt{report}, and
\texttt{article}, respectively:
%
\begin{itemize}
    \item \textbf{\texttt{hgbthesis}} (\texttt{book}):
    for bachelor's, master's, and diploma theses;
    \item \textbf{\texttt{hgbreport}} (\texttt{report}):
    for project and term reports;
    \item \textbf{\texttt{hgbarticle}} (\texttt{article}):
    for drafting journal articles.
\end{itemize}


\subsection{Class options}

\subsubsection{General options}

All document classes accept the following general options:
%
\begin{itemize}
    \item \texttt{english} or \texttt{german} 
        (select the primary document language),
    \item \texttt{smartquotes} (use smart quotes replacement),
    \item \texttt{apa} (use \texttt{apa} bibliography style instead of
    \texttt{numeric-comp}),
    \item \texttt{noUpdateCheck} (suppress check of package release date).
\end{itemize}


\subsubsection{Class-specific options}

In addition, the following class-specific options are accepted:
%
\begin{itemize}
    \item \textbf{\texttt{hgbthesis}}:\\
      \texttt{type=}%
      \{\texttt{bachelor}, \texttt{master}, \texttt{diploma}, \texttt{phd}, \texttt{internship}, \texttt{custom}\},\newline
      \texttt{language=}\{\texttt{english}, \texttt{german}\} (language in main document),%
				\footnote{Note: The \texttt{german} option implicitly loads Babel's \texttt{ngerman} setup. 
				When switching languages (see Sec.~\ref{sec:LanguageSwitchingCommands}) only \texttt{german} should be used.}
			\\
      \texttt{titlelanguage=}\{\texttt{english}, \texttt{german}\} (language on title pages, defaults to main language),\\
      \texttt{proposal} (flag to indicate proposal documents);
    \item \textbf{\texttt{hgbreport}}:\\
      \texttt{notitlepage};
    \item \textbf{\texttt{hgbarticle}}\\
      \texttt{twocolumn}.
\end{itemize}
%
For example, to start a master's thesis in German, simply place
%
\begin{LaTeXCode}[numbers=none]
\documentclass[type=master,language=german,smartquotes]{hgbthesis}
\end{LaTeXCode}
%
at the beginning of the document.

The \texttt{proposal} option is intended for a \emph{thesis proposal} (``Exposé'' or ``Proposal'' ) and is 
only effective in \emph{conjunction} with the main thesis types, \eg,
%
\begin{LaTeXCode}[numbers=none]
\documentclass[type=bachelor,language=english,titlelanguage=german,proposal]{hgbthesis}
\end{LaTeXCode}
%
The result is a short exposé that contains only one chapter. Thus,
chapter numbers are not displayed. Remove the \texttt{proposal} option to
migrate a proposal document straight to the final thesis (and restore the usual
numbering scheme).

\subsection{Thesis parameters (class \texttt{hgbthesis})}

\texttt{hgbthesis} supports several types of thesis documents. The following
parameters must be specified at the beginning of the main document:
%
\begin{itemize}
    \item \verb!\title{...}!,
    \item \verb!\subtitle{...}! (optional),
    \item \verb!\author{...}!,
    \item \verb!\programtype{...}! (optional), 
    \item \verb!\programname{...}!,
    \item \verb!\institution{...}!,
    \item \verb!\placeofstudy{...}!,
    \item \verb!\dateofsubmission{yyyy}{mm}{dd}!,
    \item \verb!\advisor[]{...}! (optional, multiple \verb!\advisor! statements allowed),
\end{itemize}
%
Note that \texttt{hgbthesis} only supports a \emph{single author} inside the
\verb!\author{...}! macro argument (commands \verb!\and! and
\verb!\thanks{...}! are deactivated)!

The command \verb!\advisor[role]{name}! accepts an optional argument to associate a \emph{role} 
with the advisor's \emph{name}, \eg, \verb!\advisor[Supervisor]{Prof.~Marie Curie, PhD}!.
Multiple advisors can be specified, for example
\begin{LaTeXCode}[numbers=none]
  \advisor[1.~Betreuerin]{Professor Frida K.~Putnik, PhD}
  \advisor[2.~Betreuer]{Franz Grillparzer, TU Wien}
  \advisor[Gutachter]{Dr.~B.\,Rutal, MIT}
\end{LaTeXCode}

\noindent
Other (optional) settings for \texttt{hgbthesis} include:
\begin{itemize}
    \item \verb!\license{cc|strict}!\newline
    Use the Creative Commons License (\texttt{cc} = default) or strict terms
      (``all rights reserved'');
    \item \verb!\logofile{xxx.pdf}!\newline
    Specifies a custom logo image for the title page, to be placed in \verb!images/xxx.pdf!
    (default is \texttt{logo.pdf}).
\end{itemize}

\section{Style files and user commands}

The package comes with a set of style (\texttt{*.sty}) files that can be used
independently of the document classes listed above:
\begin{itemize}
  \item \texttt{hgb.sty} (language, date setup, custom commands;
  \item \texttt{hgbabbrev.sty}: various abbreviation commands;
  \item \texttt{hgbalgo.sty}: additions to \texttt{algpseudocodex} package;
  \item \texttt{hgbbib.sty}: bibliography setup;
  \item \texttt{hgbdict.sty}: language dictionary functions;
  \item \texttt{hgbheadings.sty}: definition of page headers;
  \item \texttt{hgblistings.sty}: setup for code listings;
  \item \texttt{hgbmath.sty}: setup and commands for math typesetting;
  \item \texttt{hgbpdfa.sty}: setup for PDF/A generation.
\end{itemize}

\subsection{General user commands and environments (\texttt{hgb.sty})}
\label{sec:GeneralUserCommands}

\begin{itemize}
   \item \textbf{\texttt{{\bs}hgbDate}}\newline
    Outputs the package version date,
    \eg, ``\texttt{\hgbDate}''.
	\item \textbf{\texttt{{\bs}getcurrentlabel}}\newline
    Yields the most recently assigned label number.
   \item \textbf{\texttt{{\bs}calibrationbox}}\verb!{width}{height}!\newline
    Inserts a test box for checking the final print size (in millimeters).
	\item \verb!\begin{block}...\end{block}!\newline
    Dummy environment, provides a limited scope for variable/command redefinitions.
\end{itemize}


\subsection{Language switching commands (\texttt{hgb.sty})}
\label{sec:LanguageSwitchingCommands}

\begin{itemize}
	\item \verb!\SetLanguage{<language>}!\newline
    Switches to the specified language.
		This and the other language switching commands listed below
		also refresh the \texttt{smartquotes} settings (if activated).
	\item \verb!\begin{SwitchLanguage}{<language>}...\end{SwitchLanguage}!\newline
    Temporarily switches to the specified language and reverts to the current language 
		afterwards. 
   \item \verb!\begin{english}...\end{english}!\newline
    Temporarily switches to English language settings -- short for\newline
		\verb!\begin{SwitchLanguage}{english}..!.
   \item \verb!\begin{german}...\end{german}!\newline
    Temporarily switches to German language settings -- short for\newline
		\verb!\begin{SwitchLanguage}{german}..!.
\end{itemize}



\subsection{Text commands (\texttt{hgbabbrev.sty})}

\subsubsection*{Special characters:}

\begin{itemize}
    \item \textbf{\texttt{{\bs}bs}}: Inserts a backslash character (short for
    \verb!\textbackslash!).
    \item \textbf{\texttt{{\bs}obnh}}: Inserts an optional break with no
    hyphen (\eg, \verb!PlugIn{\obnh}Filter!).
\end{itemize}


\subsubsection*{German abbreviations:}

\begin{itemize}
    \item \textbf{\texttt{{\bs}bzgl}}: bzgl.
    \item \textbf{\texttt{{\bs}bzw}}: bzw.
    \item \textbf{\texttt{{\bs}ca}}: ca.
    \item \textbf{\texttt{{\bs}dah}}: d.\thinspace{}h.
    \item \textbf{\texttt{{\bs}Dah}}: D.\thinspace{}h.
    \item \textbf{\texttt{{\bs}ds}}: d.\thinspace{}sind
    \item \textbf{\texttt{{\bs}etc}}: etc.
    \item \textbf{\texttt{{\bs}evtl}}: evtl.
    \item \textbf{\texttt{{\bs}ia}}: i.\thinspace{}Allg.
    \item \textbf{\texttt{{\bs}sa}}: s.\ auch
    \item \textbf{\texttt{{\bs}so}}: s.\ oben
    \item \textbf{\texttt{{\bs}su}}: s.\ unten
    \item \textbf{\texttt{{\bs}ua}}: u.\thinspace{}a.
    \item \textbf{\texttt{{\bs}Ua}}: U.\thinspace{}a.
    \item \textbf{\texttt{{\bs}uae}}: u.\thinspace{}\"A.
    \item \textbf{\texttt{{\bs}usw}}: usw.
    \item \textbf{\texttt{{\bs}uva}}: u.\thinspace{}v.\thinspace{}a.
    \item \textbf{\texttt{{\bs}uvm}}: u.\thinspace{}v.\thinspace{}m.
    \item \textbf{\texttt{{\bs}va}}: vor allem
    \item \textbf{\texttt{{\bs}vgl}}: vgl.
    \item \textbf{\texttt{{\bs}zB}}: z.\thinspace{}B.
    \item \textbf{\texttt{{\bs}ZB}}: Zum Beispiel
\end{itemize}

\subsubsection*{English abbreviations:}

\begin{itemize}
    \item \textbf{\texttt{{\bs}ie}}: \ie
    \item \textbf{\texttt{{\bs}eg}}: \eg
    \item \textbf{\texttt{{\bs}etc}}: etc.
    \item \textbf{\texttt{{\bs}Eg}}: \Eg
    \item \textbf{\texttt{{\bs}wrt}}: \wrt
\end{itemize}

\noindent
Note that none of the above abbreviation macros ``eats'' subsequent white
space, \ie, they can be used without additional controls, as in
``\verb!\wrt what I said!'', for example.

\subsection{Bibliography commands (\texttt{hgbbib.sty})}

\begin{itemize}
    \item 
    \textbf{\texttt{{\bs}AddBibFile}}\newline
    A wrapper to \texttt{biblatex}'s \verb!\addbibresource! macro (for backward compatibility only).
    \item
    \textbf{\texttt{{\bs}MakeBibliography}}[\emph{options}]
    \newline
    Inserts the
    reference section or chapter. By default, references are automatically
    split into category subsections.%
    \footnote{Predefined reference categories are \texttt{literature},
        \texttt{avmedia}, \texttt{online} and \texttt{software}.}
    Use the option \texttt{nosplit} to produce a traditional (\ie,
    contiguous) list of references.
    \item
    \textbf{\texttt{{\bs}mcite}}%
        [\emph{text1}]\{\emph{key1}\}%
        [\emph{text2}]\{\emph{key2}\}%
        \ldots
        [\emph{textN}]\{\emph{keyN}\}%
    \newline
    Analogous to \texttt{bib\-la\-tex}'s \texttt{{\bs}cites} command%
    \footnote{%
    \url{http://mirrors.ctan.org/macros/latex/contrib/biblatex/doc/biblatex.pdf}
    (see Sec.~3.8.3)} but inserts semicolons between reference entries for
    better readability.
\end{itemize}

% \MakeBibliography ... creates a reference section split subsections (default)
% \MakeBibliography[nosplit] ... creates a one-piece reference section


\subsection{Code environments (\texttt{hgblistings.sty})}

The following types of code environments are defined:%
%
\begin{itemize}
    \item \textbf{\texttt{CCode}}: for C (ANSI),
    \item \textbf{\texttt{CppCode}}: for C++ (ISO),
    \item \textbf{\texttt{CsCode}}: for C\#,
    \item \textbf{\texttt{CssCode}}: for CSS,
    \item \textbf{\texttt{GenericCode}}: for generic code,
    \item \textbf{\texttt{HtmlCode}}: for HTML,
    \item \textbf{\texttt{JavaCode}}: for Java,
    \item \textbf{\texttt{JsCode}}: for JavaScript,
    \item \textbf{\texttt{LaTeXCode}}: for \latex,
    \item \textbf{\texttt{ObjCCode}}: for ObjectiveC,
    \item \textbf{\texttt{PhpCode}}: for PHP,
    \item \textbf{\texttt{PythonCode}}: for Python,
    \item \textbf{\texttt{Swift}}: for Swift,
    \item \textbf{\texttt{XmlCode}}: for XML.
\end{itemize}
%
\texttt{hgblistings} is based on the \texttt{listingsutf8}%
\footnote{\url{https://ctan.org/pkg/listingsutf8}}
package, thus any valid \texttt{listings}%
\footnote{\url{https://ctan.org/pkg/listings}}
option may be used; for example, the option \texttt{numbers=none} to suppress
line numbers:
%
\begin{LaTeXCode}[numbers=none]
    \begin{JavaCode}[numbers=none]
    ... // Java code comes here
    \end{JavaCode}
\end{LaTeXCode}


\subsection{Mathematical commands (\texttt{hgbmath.sty})}

\texttt{hgbmath} requires (and automatically loads) the \texttt{amsmath}%
\footnote{\url{https://ctan.org/pkg/amsmath}}
package, thus, all commands and symbols of \texttt{amsmath} are available by
default. The following \emph{additional} commands can only be used in math mode:
%
\begin{itemize}
    \item \textbf{\texttt{{\bs}Cpx}}: $\Cpx$ (complex numbers),
    \item \textbf{\texttt{{\bs}N}}: $\N$ (natural numbers),
    \item \textbf{\texttt{{\bs}R}}: $\R$ (real numbers),
    \item \textbf{\texttt{{\bs}Q}}: $\Q$ (rational numbers),
    \item \textbf{\texttt{{\bs}Z}}: $\Z$ (integer numbers).
\end{itemize}


\subsection{Algorithms (\texttt{hgbalgo.sty})}

\texttt{hgbalgo} is a stand-alone package that is based on -- and extends --
the \texttt{algorithmicx} and \texttt{algpseudocodex} packages.%
\footnote{\url{https://ctan.org/pkg/algorithmicx},
    \url{https://ctan.org/pkg/algpseudocodex}}
It fixes some (mostly indentation-related) problems, adds color, and provides
some additional commands. It also loads the \texttt{algorithm}%
\footnote{\url{https://ctan.org/pkg/algorithms}}
package, which defines a compatible float container for algorithms:
\verb!\begin{algorithm}! \verb!...! \verb!\end{algorithm}!.

\paragraph{Additional algorithm commands:}
\begin{itemize}
    \item
    \textbf{\texttt{{\bs}StateNN[<nesting>]\{<text>\}}}:
    Creates a \emph{non-numbered} statement like \texttt{algo\-rith\-micx}'s
    \verb!\Statex! command but provides controlled indentation inside nested
    constructs. The optional integer argument \verb!<nesting>! can be used to
    specify the \emph{nesting depth} to compensate for a bug in
    \texttt{algorithmicx} (the nesting level inside a block is not set
    properly before the first \verb!\State! command). Omitting the optional
    argument should give correct indentation in most situations.
    \item
    \textbf{\texttt{{\bs}Input\{<text>\}}}:
    For describing the input parameters in a procedure's preamble.
    \item
    \textbf{\texttt{{\bs}Output\{<text>\}}}:
    For describing the output values in a procedure's preamble.
    \item
    \textbf{\texttt{{\bs}Returns\{<text>\}}}:
    For describing the return values in a procedure's preamble.
\end{itemize}

\paragraph{Vertical spacing commands:}
The following commands are provided for fine-tuning the vertical spacing
between individual statements of an algorithm (the standard spacing commands
like \verb!\smallskip! \etc\ have no effect between statements):%
\footnote{Note that the standard spacing commands work \emph{between}
\texttt{procedure} and \texttt{function} blocks in the usual way.}
\begin{itemize}
    \item \textbf{{\bs}\texttt{algsmallskip}}: inserts 3pt extra space,
    \item \textbf{{\bs}\texttt{algmedskip}}: inserts 6pt extra space,
    \item \textbf{{\bs}\texttt{algbigskip}}: inserts 12pt extra space.
\end{itemize}
They are supposed to be used inside (\ie, at the end of) statements, for
example:
%
\begin{LaTeXCode}[numbers=none]
    \State $x \gets x + 1$ \algsmallskip
\end{LaTeXCode}

\paragraph{Defined algorithm colors:}
\begin{itemize}
    \item \textbf{\texttt{AlgKeywordColor}} (for algorithm keywords),
    \item \textbf{\texttt{AlgProcedureColor}} (for procedure and function
    names).
\end{itemize}
These colors can be redefined at any time (see the \texttt{xcolor}%
\footnote{\url{https://ctan.org/pkg/xcolor}}
package), \eg, by
\begin{LaTeXCode}[numbers=none]
    \definecolor{AlgKeywordColor}{named}{black}
    \definecolor{AlgProcedureColor}{rgb}{0.0, 0.5, 0.0}     % dark green
\end{LaTeXCode}


\section{PDF/A generation (\texttt{hgbpdfa.sty})}

Package \texttt{hgbpdfa} contains definitions for generating PDF/A-compliant 
(PDF/A-2b) output files. It is based on the \texttt{pdfmanagement-testphase}
package (requires version 0.95s or higher) and must be loaded before the
\verb!\documentclass! statement in the main document, for example, by
\begin{LaTeXCode}[numbers=none]
\RequirePackage{hgbpdfa}
\end{LaTeXCode}
If omitted, a plain PDF (non-PDF/A-compliant) file is generated.


\section{Customizing thesis title pages}

The content and structure of the title pages generated for the various thesis types
(\texttt{bachelor}, \texttt{master} \etc) may be customized to meet the specific
requirements of different institutions or departments.
Thus customization is usually done at the institutional level and not
by the individual author (student).

\subsection{Standard title page setups (\emph{themes})}

The various front page arrangements are called \emph{themes} in the following.
Each theme is identified by a unique \texttt{themeID} and associated with a 
particular style (\latex\ package) file named \verb!hgbtheme-<themeID>.sty!.
For example, the \texttt{default} theme  is defined by file
\begin{itemize}
  \item[] \texttt{hgbtheme-default.sty},
  %\item[] \texttt{hgbtheme-custom.sty}.
\end{itemize}
%
Additional resources required by a theme (such as graphics files) must be named with the 
complete \emph{theme name} (\verb!hgbtheme-<themeID>!) as a prefix, for example,%
\footnote{This is to avoid file name conflicts since themes are part of the 
CTAN distribution and thus all theme-related files will eventually end up in a single, 
flat directory.}
\begin{itemize}
	\item[] \texttt{hgbtheme-default-logo.pdf}
\end{itemize}
%
which contains the logo graphics used by that theme.

To \emph{use} a specific theme, option \verb!theme=<themeID>! is added to the 
\verb!\documentclass! command, for example,
\begin{itemize}
	\item[] \verb!\documentclass[theme=default,...]{hgbthesis}!
\end{itemize}
%All theme sources are stored in `/dev/themes` (perhaps with individual subdirectories?).
%\hgb@UseTheme{default}%


All these ``inherit'' from a parent theme file
%
\begin{itemize}
  \item[] \texttt{hgbtheme-thesis.sty},
\end{itemize}
%
which contains the bulk of the definitions, while derived themes typically contain only 
a small number of specific settings.


\subsection{Customizing themes}

To customize the title page setup to their needs, authors (or administrators) should 
\emph{not} modify any of the standard theme files, since these may not be local but loaded
from a package distribution.
There are two recommended ways instead:

\subsubsection{Option 1: Adapt \texttt{hgbtheme-custom.sty}}

Copy the theme file\texttt{hgbtheme-custom.sty} (which is part of this distribution) to the 
\emph{main document directory} (if not there already) and open it in your LaTeX editor.
It defines \texttt{hgbtheme-default.sty} as the parent theme by the initial command
\begin{LaTeXCode}[numbers=none]
\hgb@UseTheme{default}
\end{LaTeXCode}
%
All commands defined in the parent theme are visible and may be redefined,
typically by \verb!\renewcommand! (see Sec.\ \ref{sec:ThemeStyleCommands}
for available commands).
To activate the associated theme simply use \texttt{custom} as the theme ID,
\ie,
%
\begin{LaTeXCode}[numbers=none]
\documentclass[theme=custom,...]{hgbthesis}
\end{LaTeXCode}
%
This is the simplest approach if only a \emph{single} custom theme is needed.

\subsubsection{Option 2: Create multiple custom themes}

Thesis administrator may find it useful to define \emph{multiple} custom themes for their
institution or department(s).
For this purpose, simply copy file \texttt{hgbtheme-default.sty} or \texttt{hgbtheme-custom.sty}
to a new file, \eg, \texttt{hgbtheme-physics.sty},%
\footnote{To be placed in the main document directory. Note the naming conventions!}
and modify it accordingly. The associated theme can then be activated by
%
\begin{LaTeXCode}[numbers=none]
\documentclass[theme=physics,...]{hgbthesis}
\end{LaTeXCode}
%
An error will be raised if the associated \texttt{.sty} file cannot be found.



\subsection{Commands and variables available to custom theme styles}
\label{sec:ThemeStyleCommands}

Any theme style must at least \emph{redefine} the
\begin{itemize}
  \item[] \verb!\hgb@MakeFrontPages{..}!
\end{itemize}
command, which is pre-defined (as throwing an error) in \texttt{hgbthesis.cls}
and invoked by \verb!\maketitle!.
If \texttt{default} is used as the parent theme, \verb!\hgb@MakeFrontPages! 
is already set up properly and redefinition is optional, \ie,
%
\begin{LaTeXCode}[numbers=none]
\hgb@UseTheme{default}		% parent theme
\renewcommand{\hgb@MakeFrontPages}{..}	% optional
\end{LaTeXCode}
%

\noindent
Moreover, the following macros and variables are assured to be available for defining custom themes
(see \texttt{hgbtheme-default.sty} for their typical usage).
These are defined in \texttt{hgbthesis.cls} and should \emph{not} be redefined:
%
\begin{itemize}
  \item[] \verb!\hgb@Author!
  \item[] \verb!\hgb@Title!
  \item[] \verb!\hgb@SubTitle!
  \item[] \verb!\hgb@Institution!
  \item[] \verb!\hgb@ProgramType!
  \item[] \verb!\hgb@ProgramName!
  \item[] \verb!\hgb@PlaceOfStudy!
\medskip
	\item[] \verb!\hgb@ThesisName!
	\item[] \verb!\hgb@ProposalName!
\medskip
  \item[] \verb!\hgb@AdvisorCount!
  \item[] \verb!\hgb@getAdvisorRole{<number>}!
  \item[] \verb!\hgb@getAdvisorName{<number>}!
\medskip
  \item[] \verb!\hgb@MainLanguage!
  \item[] \verb!\hgb@TitleLanguage!
  \item[] \verb!\hgb@TitlePageFont!
  %\item[] \verb!\hgb@LogoFile!
\medskip 
  %\item[] \verb!\hgb@Advisor!
  \item[] \verb!\hgb@SubmissionYear!
  \item[] \verb!\hgb@SubmissionMonth!
  \item[] \verb!\hgb@SubmissionDay!
  \item[] \verb!\hgb@GetMonthName{<language>}{<monthnumber>}!
\medskip
	\item[] \verb!\hgb@License!
\medskip
  \item[] \verb!hgb@IsProposal! (\emph{boolean}, without \verb!\!)
\end{itemize}

\noindent
Class \texttt{hgbthesis.cls} also defines a special \emph{hook}
\begin{itemize}
\item[] \verb!hgb@InitThemeHook!
\end{itemize}
for adding custom initialization code for the current theme in the form
%
\begin{LaTeXCode}[numbers=none]
\AddToHook{hgb@InitThemeHook}{<initialization code>},
\end{LaTeXCode}
%
typically placed at the beginning of the theme file. The collected code for
this hook is executed immediately before \verb!\maketitle!.


In addition, theme \texttt{hgbtheme-default.sty} defines the following
commands for generating individual title pages:
%
\begin{itemize}
  \item[] \verb!\hgb@MakeTitlePage!
  \item[] \verb!\hgb@MakeAdvisorPage!
  \item[] \verb!\hgb@MakeCopyrightPage!
  \item[] \verb!\hgb@MakeDeclarationPage!
\end{itemize}
%
Each of these may be redefined by inheriting themes.

\section{Package dependencies}

\begin{sloppypar}
The \texttt{hagenberg-thesis} package builds on the following \latex\ packages:\newline
\texttt{abstract}, 
\texttt{algorithm}, 
\texttt{algorithmicx}, 
\texttt{algpseudocodex}, 
\texttt{amsbsy}, 
\texttt{amsfonts}, 
\texttt{amsmath}, 
\texttt{amssymb}, 
\texttt{babel}, 
\texttt{biblatex}, 
\texttt{breakurl}, 
\texttt{caption}, 
\texttt{cmap}, 
\texttt{csquotes}, 
\texttt{datetime2}, 
\texttt{enumitem}, 
\texttt{epstopdf}, 
\texttt{exscale}, 
\texttt{fancyhdr}, 
\texttt{float}, 
\texttt{fontenc},
\texttt{forloop},
\texttt{geometry}, 
\texttt{graphicx}, 
\texttt{hypcap}, 
\texttt{hyperref}, 
\texttt{ifpdf}, 
\texttt{inputenc}, 
\texttt{lengthconvert},
\texttt{listingsutf8}, 
\texttt{lmodern},
\texttt{marvosym}, 
\texttt{moreverb}, 
\texttt{overpic}, 
\texttt{pdfmanagement-testphase}, 
\texttt{pdfpages}, 
\texttt{pict2e}, 
\texttt{subdepth}, 
\texttt{titlesec}, 
\texttt{titling},
\texttt{tocbasic},
\texttt{url}, 
\texttt{upquote}, 
\texttt{verbatim}, 
\texttt{xcolor}, 
\texttt{xifthen},
\texttt{xstring},
\texttt{xspace}.
\end{sloppypar}

\end{document}
