%% A simple template for a term report using the Hagenberg setup
%% based on the standard LaTeX 'report' class
%% äöüÄÖÜß  <-- no German Umlauts here? Use an UTF-8 compatible editor!

%%% Magic comments for setting the correct parameters in compatible IDEs
% !TeX encoding = utf8
% !TeX program = pdflatex 
% !TeX spellcheck = en_US
% !BIB program = biber

\documentclass[notitlepage,english]{hgbreport}

\RequirePackage[utf8]{inputenc}		% remove when using lualatex oder xelatex!

\graphicspath{{images/}}  % where are the images?
\bibliography{references}  % requires file 'references.bib'

%%%----------------------------------------------------------
\author{Peter A.\ Wiseguy}
\title{CS799 Ridiculously Advanced Systems\\ % the name of the course or project
			Project Report}	% or "Term Report"
\date{\today}
%%%----------------------------------------------------------


%%%----------------------------------------------------------
\begin{document}
%%%----------------------------------------------------------

\maketitle

\begin{abstract}\noindent
This document is a simple template for a typical term or semester paper (lab/course report, 
``Übungsbericht'', \etc) based on the \textsf{HagenbergThesis} \latex package.%
\footnote{See \url{https://github.com/Digital-Media/HagenbergThesis} for the most current version.
This repository also provides a good introduction and useful hints for authoring academic texts with LaTeX.}
The structure and chapter titles have been formulated to provide a good starting point
for a typical \emph{project report}.
This document uses the custom class \textsf{hgbreport} which is based on \latex's standard \textsf{report} 
document class with \texttt{chapter} as the top structuring element. 
If you wish to write this report in German you should substitute the line
\begin{quote}
 \verb!\documentclass[english]{hgbreport}! 
\end{quote}
at the top of this document by
\begin{quote}
 \verb!\documentclass[german]{hgbreport}!.
\end{quote}
Also, you may want to place the text of the individual chapters in separate files and 
include them using \verb!\include{..}!.

\bigskip
\noindent
Use the abstract to provide a short summary of the contents in the document.
\end{abstract}


%%%----------------------------------------------------------
\tableofcontents
%%%----------------------------------------------------------



%%%----------------------------------------------------------
\chapter{Aims and Context}
%%%----------------------------------------------------------

Describe the initial goals and situation that lead to this project, requirements, as well as
references to related work (e.g., \cite{Higham1998}).



%%%----------------------------------------------------------
\chapter{Project Details}
%%%----------------------------------------------------------

Describe important project steps, e.g., rationale of the chosen architecture or technology stack, 
design decisions, algorithms used, interesting challenges faced on the way, lessons 
learned \etc




%%%----------------------------------------------------------     
\chapter{System Documentation}
%%%----------------------------------------------------------

Give a well-structured description of the architecture and the technical design
o your implementation,
with sufficient granularity to enable an external person to
continue working on the project.


%%%----------------------------------------------------------
\chapter{Summary}
%%%----------------------------------------------------------

Give a concise (and honest) summary of what has been accomplished and what not. 
Point out issues that may warrant further investigation.


\appendix %%%-----------------------------------------------


%%%----------------------------------------------------------
\chapter{Supplementary Materials}
%%%----------------------------------------------------------

The appendix is a good place to attach a user guide, screenshots, installation instructions etc.
Add a separate chapter for each major item.


%%%----------------------------------------------------------
\MakeBibliography[nosplit]
%%%----------------------------------------------------------


\end{document}
